Study 2 Analysis

### 2024-10-09 study2.qmd
update
add calculator and education plots & tables; add % of sbjs matching target plots

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here, haven,tibble,ggplot2,gt,ggh4x,lme4)

@@

s2 <- haven::read_sav("data/Frequency & Probability Study 2 - 3-24-19.sav")
s2 <- s2 |> mutate(id = as.numeric(format(EndDate, "%H%M%S"))) |> relocate(id)

s2 <- s2 |> 
  mutate(
    state1 = case_when(
      COLORADO == 1 ~ "Colorado",
      MASSACHUSETTS == 1 ~ "Massachusetts",
      CALIFORNIA == 1 ~ "California",
      TEXAS == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      COLORADO == 2 ~ "Colorado",
      MASSACHUSETTS == 2 ~ "Massachusetts",
      CALIFORNIA == 2 ~ "California",
      TEXAS == 2 ~ "Texas",
      TRUE ~ NA_character_
    )
  ) |> 
  mutate(refClass = case_when(
  REFERENCECLASS == 1 ~ "USD",
  REFERENCECLASS == 2 ~ "kWh",
  REFERENCECLASS == 3 ~ "Percentage"
)) |> 
  relocate(refClass,state1,state2, .after = id)

s2_b1 <- s2 |> select(id,state1,refClass,TENPERCENT,ROUNDED, ACFAMILY1:TOTALAPTWO1) |> 
  mutate( 
    block = 1,
    state=state1,
  pct_goal=case_when (TENPERCENT == 1 ~ "10%", TENPERCENT == 2 ~ "15%"), 
  rounded = case_when(ROUNDED == 1 ~ "Rounded", ROUNDED == 2 ~ "Not Rounded")
  ) |> relocate(id,block,refClass,pct_goal,rounded,state)  |> 
  select(-TENPERCENT,-ROUNDED, -state1)

s2_b2 <- s2 |> select(id,state2,refClass,TENPERCENT,ROUNDED, ACFAMILY2:TOTALAPTWO2) |> 
  mutate( 
    block = 2,
    state=state2,
  pct_goal=case_when (TENPERCENT == 1 ~ "15%", TENPERCENT == 2 ~ "10%"), 
  rounded = case_when(ROUNDED == 1 ~ "Not Rounded", ROUNDED == 2 ~ "Rounded")
  ) |> relocate(id,block,refClass,pct_goal,rounded,state) |> 
  select(-TENPERCENT,-ROUNDED, -state2)

@@

# compute total kWh reduction, and total % reduction 
sum1 <- s2_long |> 
  filter(appliance !="TOTAL") |> 
  group_by(id,refClass,state,pct_goal,rounded,block,plan) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif))

@@

outliers <- sum1 |>  ungroup() |> group_by(state) |> mutate(change_mean=mean(pct_change,na.rm=TRUE),change_sd=sd(pct_change,na.rm=TRUE),
                                     z_score=(pct_change-change_mean)/change_sd,
                                     is_outlier=abs(z_score)>2.5)

@@

@@

>>>>

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here, haven,tibble,ggplot2,gt,ggh4x,lme4,flextable)

@@

s2 <- haven::read_sav("data/Frequency & Probability Study 2 - 3-24-19.sav")
s2 <- s2 |> mutate(id = as.numeric(format(EndDate, "%H%M%S"))) |> relocate(id)
s2 <- s2 |> 
  mutate(
    state1 = case_when(
      COLORADO == 1 ~ "Colorado",
      MASSACHUSETTS == 1 ~ "Massachusetts",
      CALIFORNIA == 1 ~ "California",
      TEXAS == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      COLORADO == 2 ~ "Colorado",
      MASSACHUSETTS == 2 ~ "Massachusetts",
      CALIFORNIA == 2 ~ "California",
      TEXAS == 2 ~ "Texas",
      TRUE ~ NA_character_
    )
  ) |> 
  mutate(refClass = case_when(
  REFERENCECLASS == 1 ~ "USD",
  REFERENCECLASS == 2 ~ "kWh",
  REFERENCECLASS == 3 ~ "Percentage"
)) |> 
  rename(calc=MATH01) |>  # replace NA's in calc with 0's
  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
  relocate(refClass,state1,state2,calc, .after = id)

# recode DEMO4 variable based on this mapping: 
# 1 = "Some schooling, but no diploma or degree" 
# 2 = "Highschool diploma or GED" 
# 3 = "Some college" 
# 4 = "College degree" 
# 5 = "Some graduate school" 
# 6 = "Graduate degree" 
# 7 = "Choose not to answer"

s2 <- s2 |> rename(edu=DEM04) |>
  mutate(edu = factor(case_when(
    edu == '1' ~ "Some schooling, but no diploma or degree",
    edu == 2 ~ "Highschool diploma or GED",
    edu == 3 ~ "Some college",
    edu == 4 ~ "College degree",
    edu == 5 ~ "Some graduate school",
    edu == 6 ~ "Graduate degree",
    edu == 7 ~ "Choose not to answer",
    TRUE ~ NA_character_
  ))) |> relocate(edu, .after = id)

s2_b1 <- s2 |> select(id,state1,refClass,TENPERCENT,ROUNDED, calc, edu,ACFAMILY1:TOTALAPTWO1) |> 
  mutate( 
    block = 1,
    state=state1,
  pct_goal=case_when (TENPERCENT == 1 ~ "10%", TENPERCENT == 2 ~ "15%"), 
  pct=case_when (TENPERCENT == 1 ~ .10, TENPERCENT == 2 ~ .15),
  rounded = case_when(ROUNDED == 1 ~ "Rounded", ROUNDED == 2 ~ "Not Rounded")
  ) |> relocate(id,block,refClass,pct_goal,rounded,state)  |> 
  select(-TENPERCENT,-ROUNDED, -state1)

s2_b2 <- s2 |> select(id,state2,refClass,TENPERCENT,ROUNDED, calc,edu,ACFAMILY2:TOTALAPTWO2) |> 
  mutate( 
    block = 2,
    state=state2,
  pct_goal=case_when (TENPERCENT == 1 ~ "15%", TENPERCENT == 2 ~ "10%"), 
  pct=case_when (TENPERCENT == 1 ~ .10, TENPERCENT == 2 ~ .15),
  rounded = case_when(ROUNDED == 1 ~ "Not Rounded", ROUNDED == 2 ~ "Rounded")
  ) |> relocate(id,block,refClass,pct_goal,rounded,state) |> 
  select(-TENPERCENT,-ROUNDED, -state2)

@@

# compute total kWh reduction, and total % reduction 
sum1 <- s2_long |> 
  filter(appliance !="TOTAL") |> 
  group_by(id,refClass,state,pct_goal,pct,rounded,block,plan,calc,edu) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif))

@@

outliers <- sum1 |>  ungroup() |> group_by(state) |> mutate(change_mean=mean(pct_change,na.rm=TRUE),change_sd=sd(pct_change,na.rm=TRUE),
                                     z_score=(pct_change-change_mean)/change_sd,
                                     is_outlier=abs(z_score)>2.75)

@@

## Looking at % of subjects who hit the target reduction

# Step 1: Calculate whether participants matched the target pct_goal
sum1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id, refClass, state, pct_goal, pct, rounded, block, plan, calc, edu) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct))

# Step 2: Summarize the percentage of participants meeting the goal for each condition
summary_pct <- sum1 |>
  group_by(refClass, rounded, pct_goal) |>
  summarise(
    matched_count = sum(matched_goal),
    total_count = n(),
    pct_matched = matched_count / total_count * 100,
    .groups = "drop"
  )

# Step 3: Create plots for the percentage of subjects in each condition who matched the target pct
ggplot(summary_pct, aes(x = refClass, y = pct_matched, fill = rounded)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  facet_wrap(~ pct_goal) +
  labs(
    title = "Percentage of Participants Matching Target Pct by Condition",
    x = "Reference Class",
    y = "Percentage Matched (%)",
    fill = "Rounding"
  ) +
  theme_minimal()

summary_pct |>
  gt() |>
  tab_header(
    title = md("**Percentage of Participants Matching Target Pct by Condition**"),
    subtitle = "Grouped by Reference Class, Rounding, and Pct Goal"
  ) |>
  fmt_number(
    columns = c(pct_matched),
    decimals = 2
  ) |>
  cols_label(
    refClass = "Reference Class",
    rounded = "Rounding",
    pct_goal = "Pct Goal",
    matched_count = "Matched Count",
    total_count = "Total Count",
    pct_matched = "Percentage Matched (%)"
  ) |>
  cols_align(
    align = "center",
    columns = everything()
  ) |>
  tab_spanner(
    label = "Participants",
    columns = c(matched_count, total_count, pct_matched)
  )

```{r}
#| layout-ncol: 2
#| fig.height: 10
#| fig.width: 12

# Step 1: Calculate whether participants matched the target pct_goal within different tolerances
sum1a <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id, refClass, state, pct_goal, pct, rounded, block, plan, calc, edu) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(
    target_pct = as.numeric(sub("%", "", pct_goal)) / 100,
    match_level = case_when(
      abs(pct_change - target_pct) <= 0.001 ~ "Exact Match",
      abs(pct_change - target_pct) <= 0.05 ~ "Within 5%",
      abs(pct_change - target_pct) <= 0.10 ~ "Within 10%",
      abs(pct_change - target_pct) <= 0.20 ~ "Within 20%",
      TRUE ~ "Over 20%"
    ),
    match_level = factor(match_level, levels = c("Exact Match", "Within 5%", "Within 10%", "Within 20%", "Over 20%"))
  )

# Step 2: Summarize the percentage of participants in each match level for each condition
summary_pct <- sum1a |>
  group_by(refClass, rounded, pct_goal, match_level) |>
  summarise(count = n(), .groups = "drop") |>
  group_by(refClass, rounded, pct_goal) |>
  mutate(
    total_count = sum(count),
    percentage = count / total_count * 100
  ) |>
  ungroup()

# Step 3: Create plot
ggplot(summary_pct, aes(x = interaction(refClass, rounded), y = percentage, fill = match_level)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~ pct_goal, scales = "free_x") +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Percentage of Participants by Match Level",
    subtitle = "Grouped by Reference Class, Rounding, and Pct Goal",
    x = "Reference Class and Rounding",
    y = "Percentage (%)",
    fill = "Match Level"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(summary_pct, aes(x = refClass, y = percentage, fill = match_level)) +
  geom_bar(stat = "identity",position=position_dodge()) +
  facet_wrap(rounded ~ pct_goal, scales = "free_x") +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Percentage of Participants by Match Level",
    subtitle = "Grouped by Reference Class, Rounding, and Pct Goal",
    x = "Reference Class and Rounding",
    y = "Percentage (%)",
    fill = "Match Level"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
## Cumulative plot

```{r}
#| layout-ncol: 2
#| fig.height: 11
#| fig.width: 12

# Step 1: Calculate match levels with cumulative groups
sum1c <- s2_long |>
  filter(appliance != "TOTAL") |>
  group_by(id, refClass, state, pct_goal, pct, rounded, block, plan, calc, edu) |>
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(
    target_pct = as.numeric(sub("%", "", pct_goal)) / 100,
    match_level = case_when(
      abs(pct_change - target_pct) <= 0.001 ~ "Exact Match",
      abs(pct_change - target_pct) <= 0.05 ~ "Within 5%",
      abs(pct_change - target_pct) <= 0.10 ~ "Within 10%",
      abs(pct_change - target_pct) <= 0.20 ~ "Within 20%",
      TRUE ~ "Over 20%"
    ),
    match_level = factor(
      match_level, 
      levels = c("Exact Match", "Within 5%", "Within 10%", "Within 20%", "Over 20%")
    )
  )

# Step 2: Create cumulative groups for each match level
summary_pct <- sum1c |>
  group_by(refClass, rounded, pct_goal, match_level) |>
  summarise(count = n(), .groups = "drop") |>
  group_by(refClass, rounded, pct_goal) |>
  mutate(
    total_count = sum(count),
    cumulative_count = cumsum(count),
    cumulative_percentage = cumulative_count / total_count * 100
  ) |>
  ungroup()

# Step 3: Create a plot for cumulative percentages
ggplot(summary_pct, aes(x = refClass, y = cumulative_percentage, fill = match_level)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  facet_wrap(pct_goal~rounded, scales = "free_x") +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Cumulative Percentage of Participants by Match Level",
    subtitle = "Grouped by Reference Class, Rounding, and Pct Goal",
    x = "Reference Class",
    y = "Cumulative Percentage (%)",
    fill = "Match Level"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(summary_pct, aes(x = refClass, y = cumulative_percentage, fill = match_level)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  facet_wrap(~rounded, scales = "free_x") +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Cumulative Percentage of Participants by Match Level",
    subtitle = "Grouped by Reference Class, Rounding, and Pct Goal",
    x = "Reference Class",
    y = "Cumulative Percentage (%)",
    fill = "Match Level"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(summary_pct, aes(x = refClass, y = cumulative_percentage, fill = match_level)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Cumulative Percentage of Participants by Match Level",
    subtitle = "Grouped by Reference Class, Rounding, and Pct Goal",
    x = "Reference Class",
    y = "Cumulative Percentage (%)",
    fill = "Match Level"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(summary_pct, aes(x = rounded, y = cumulative_percentage, fill = match_level)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Cumulative Percentage of Participants by Match Level",
    subtitle = "Grouped by Reference Class, Rounding, and Pct Goal",
    x = "Rounding",
    y = "Cumulative Percentage (%)",
    fill = "Match Level"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

## Calculator and education

# table with only education level counts, just use kable
s2 |> 
  group_by(edu) |> 
  summarise(n = n(), .groups = "drop") |>
  knitr::kable(caption = "Count of Items by Education Level", format = "html")

# table with only calculator counts

s2 |> 
  mutate(calc = if_else(calc == 0, "No Calculator", "Used Calculator")) |> 
  group_by(calc) |> 
  summarise(n = n(), .groups = "drop") |>
  knitr::kable(caption = "Count of Items by Calculator Use", format = "html")

# s2 |> 
#   mutate(rounded = case_when(ROUNDED == 1 ~ "Not Rounded", ROUNDED == 2 ~ "Rounded")) |> 
#   group_by(calc, refClass, rounded) |> 
#   summarise(n = n(), .groups = "drop") |>
#   group_by(calc, refClass) |>
#   mutate(count_stats = as.character(n)) |>
#   ungroup() |>
#   tabulator(
#     rows = c("calc", "refClass"),
#     columns = "rounded",
#     `Count` = as_paragraph(count_stats)
#   ) |>
#   as_flextable() |> 
#   align(align = "left", part = "all") |>
#   autofit() |>
#   set_caption(
#     caption = "Count of Items by Calculation Type, Reference Class, and Rounded Value",
#     style = "Table Caption"
#   )

s2 |> 
  mutate(
    rounded = if_else(ROUNDED == 1, "Not Rounded", "Rounded"),
    calc = if_else(calc == 0, "No Calculator", "Used Calculator")
  ) |> 
  count(calc, refClass, rounded) |>
  pivot_wider(
    names_from = c(calc, rounded),
    values_from = n,
    values_fill = 0
  ) |>
  gt() |>
  tab_header(
    title = md("**Count of Items by Calculator Use, Reference Class, and Rounding**")
  ) |>
  cols_align(align = "left") |>
  tab_spanner_delim(delim = "_") |>
  cols_label(
    `No Calculator_Not Rounded` = "Not Rounded",
    `No Calculator_Rounded` = "Rounded",
    `Used Calculator_Not Rounded` = "Not Rounded",
    `Used Calculator_Rounded` = "Rounded"
  )

# table with education level counts

s2 |>
  group_by(edu, refClass) |>
  summarise(n = n(), .groups = "drop") |>
  pivot_wider(names_from = refClass, values_from = n, values_fill = 0) |>
  gt(rowname_col = "edu") |>
  tab_header(title = md("**Count of Items by Education Level and Reference Class**")) |>
  cols_align(align = "left", columns = everything())

# table with calculator and education level counts
s2 |>
  mutate(calc = if_else(calc == 0, "No Calculator", "Used Calculator")) |> 
  group_by(calc, edu) |>
  summarise(n = n(), .groups = "drop") |>
  pivot_wider(names_from = edu, values_from = n, values_fill = 0) |>
  gt(rowname_col = "calc") |>
  tab_header(title = md("**Count of Items by Calculator Use and Education Level**")) |>
  cols_align(align = "left", columns = everything())

sum1 |> ggplot(aes(x = calc, y = pct_change, fill = calc)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  labs(
    title = "Effect of Calculator Use",
    x = "Calculator Use",
    y = "% Change",
    fill = "Calculator Use"
  )

## education

sum1 |> ggplot(aes(x = edu, y = pct_change, fill = edu)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  labs(
    title = "Effect of Education",
    x = "Education",
    y = "% Change",
    fill = "Education"
  )
